---
title: "Choosing WCC Parameters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing WCC Parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Selecting hyperparameters is often the most challenging aspect of running a Windowed Cross-Correlation (WCC) analysis. Because behavioral time series vary wildly in their speed and structure, there is no universal set of defaults. A mother-infant gaze interaction unfolds on a completely different timescale than a high-speed athletic exchange.

The `bsync` package provides the `suggest_wcc_params()` function to help you ground your parameter selection in the physical properties of your data and the theoretical boundaries of the behavior you are studying.

## A Real-World Example: Conversational Head Nods

Imagine we are studying interpersonal synchrony during a conversation. We have video tracking data sampled at 30 Hz (30 frames per second), and we are specifically interested in how Person B nods in response to Person A's nods.

To use `suggest_wcc_params()`, we need to define three theoretical values: the typical duration of the event, the maximum plausible delay, and the desired window overlap.

```{r example-setup}
library(bsync)

# 30 Hz sampling rate
fs <- 30

suggest_wcc_params(
  sample_rate = fs,
  event_duration_sec = 1,
  max_delay_sec = 2,
  overlap_pct = 0.5
)
```

### 1. Event Duration (`event_duration_sec`)

The first step is to estimate how long a single instance of your target behavior takes to complete. In our example, a typical conversational head nod takes approximately 1 second to begin, peak, and return to rest.

The helper function uses this value to calculate your `window_size`. It intentionally suggests a window that is 4 times the duration of the event (in our case, 4 seconds, which translates to 120 frames at 30 Hz).

This 4x multiplier serves two critical purposes:

1. **Statistical Stability:** Pearson correlations become highly unstable with very small sample sizes. A wider window ensures there are enough data points to reliably calculate a correlation coefficient.
2. **Contextual Capture:** To identify that two signals are correlated, the window must be large enough to capture the "quiet" periods before and after the movement. If the window only captured the peak of the nod, the data would appear as a flat line, making correlation impossible.

### 2. Maximum Delay (`max_delay_sec`)

Next, we must consider the biological or psychological limits of the interaction. If Person A nods, how long does Person B have to respond before their nod is considered a new, independent action rather than a reaction?

In conversational dynamics, a response delay of more than 2 seconds is often considered a completely new initiation. We set `max_delay_sec = 2`. The helper function converts this directly into frames to set your `lag_max` (60 frames).

It is important to note that `lag_max` mathematically cannot exceed half of your `window_size`. If you shift the two signals too far apart relative to the window, the overlapping region where correlation can be calculated becomes dangerously small. The helper function will automatically cap your lag if it detects this imbalance.

### 3. Window Overlap (`overlap_pct`)

The `overlap_pct` parameter dictates the temporal resolution of your final elapsed time axis. This controls how far the window shifts forward after calculating each correlation landscape.

This parameter is a direct trade-off between computational speed and visual smoothness:

- **0% Overlap:** The window jumps completely forward (e.g., 4 seconds at a time). This is extremely fast to compute but results in a blocky, low-resolution plot that might skip over rapid shifts in synchrony.
- **50% to 75% Overlap:** A balanced choice for exploratory data analysis. The window slides forward incrementally, providing a smooth curve of correlation changes while keeping processing times low.
- **High Overlap (Increment of 1):** The window moves forward by exactly 1 frame at a time. This provides the highest possible resolution and the smoothest peak tracking. Because `bsync` is optimized in C++, many modern machines can handle an increment of 1 for typical datasets, making it ideal for final publication plots.

By setting `overlap_pct = 0.5`, the helper function suggests a `window_increment` of 60 frames. This means our 120-frame window will shift forward by 60 frames on each step, ensuring continuous coverage of the conversation.

## Putting it into Practice

Once you have reviewed the suggested parameters, you can plug them directly into the core `wcc()` function to run your analysis with confidence.

```{r example-run, eval=FALSE}
# Using the suggested parameters from above
results <- wcc(
  x = df$person_A_nod,
  y = df$person_B_nod,
  window_size = 120,
  lag_max = 60,
  window_increment = 60,
  lag_increment = 1
)
```
