---
title: "Windowed Cross-Correlation Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Windowed Cross-Correlation Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Analyzing Interpersonal Synchrony with Windowed Cross-Correlation

This vignette walks through a complete Windowed Cross-Correlation (WCC) analysis using the `bsync` package. WCC is highly effective for quantifying interpersonal synchrony because it accommodates non-stationary relationships. Unlike global cross-correlation, WCC captures how synchronization ebbs and flows over time.

We will cover data simulation, calculating WCC, surrogate testing for statistical significance, peak picking, and visualization. For guidance on selecting the appropriate window sizes and lag parameters, please see the `suggest_wcc_params()` vignette.

## 1. Simulating Realistic Dyadic Data

To demonstrate the workflow, we will simulate a realistic interaction between two participants (Person A and Person B) captured at 30 Hz. We will generate smooth continuous data to simulate bodily motion.

In this scenario, Person A leads the interaction by 15 frames (0.5 seconds) for the first half of the task. In the second half, the dynamic flips and Person B leads by 15 frames.

```{r setup-data}
library(bsync)
library(dplyr)

set.seed(2026)

# Simulation parameters
fs <- 30
n_frames <- 1800 # 60 seconds of data

# Generate a smooth base signal using a wide moving average on white noise
# A 90-frame window ensures high autocorrelation at our 15-frame shift
raw_noise <- rnorm(n_frames + 300)
base_signal <- stats::filter(raw_noise, rep(1/90, 90), circular = TRUE)
base_signal <- as.numeric(base_signal)

person_A <- numeric(n_frames)
person_B <- numeric(n_frames)

# Create the shifting leader-follower dynamic
for (i in 1:n_frames) {
  if (i <= n_frames / 2) {
    # First half: Person B follows Person A by 15 frames
    idx_A <- 150 + i
    idx_B <- 150 + i - 15
  } else {
    # Second half: Person A follows Person B by 15 frames
    idx_A <- 150 + i - 15
    idx_B <- 150 + i
  }

  # Add slight independent noise to mimic realistic measurement error
  person_A[i] <- base_signal[idx_A] + rnorm(1, sd = 0.05)
  person_B[i] <- base_signal[idx_B] + rnorm(1, sd = 0.05)
}

dyad_data <- data.frame(
  time = seq(0, by = 1/fs, length.out = n_frames),
  person_A = person_A,
  person_B = person_B
)
```

By adding a custom simulation here, we avoid relying on highly simplified data provided in the package's raw data files.

## 2. Calculating Windowed Cross-Correlation

With our data ready, we can run the primary `wcc()` function. This function slides a window across the time series and calculates the cross-correlation at various lags within each window.

We will use a window size of 90 frames (3 seconds) and a maximum lag of 45 frames (1.5 seconds). We will also increment the window by 30 frames (1 second) to allow for overlap and smooth transitions.

```{r calculate-wcc}
wcc_results <- wcc(
  x = dyad_data$person_A,
  y = dyad_data$person_B,
  window_size = 90,
  lag_max = 45,
  window_increment = 30,
  lag_increment = 1
)

# View a summary of the results
summary(wcc_results)
```

The `wcc()` function returns a list object of class `wcc_res` containing the results data frame, the overall Fisher's Z score, and the input settings.

## 3. Surrogate Testing for Significance

Time series data are inherently autocorrelated. Because of this autocorrelation, high cross-correlation values can sometimes occur purely by chance. To test if the synchronization we observed is meaningful, we generate a null distribution using the circular shift method.

The `wcc_surrogate()` function handles this by shifting one time series relative to the other, destroying the true synchronous relationship while preserving the autocorrelation of the individual signals.

**Note on Permutations:** We use `n_surrogates = 100` here for speed during exploratory analysis. To achieve stable and trustworthy p-values for publication, it is recommended to run at least 1,000 to 10,000 permutations.

```{r surrogate-testing}
surrogate_results <- wcc_surrogate(
  x = dyad_data$person_A,
  y = dyad_data$person_B,
  window_size = 90,
  lag_max = 45,
  window_increment = 30,
  lag_increment = 1,
  n_surrogates = 100
)

print(surrogate_results)
```

The output gives us an empirical p-value by calculating the proportion of surrogate Fisher's Z scores that meet or exceed our observed Fisher's Z. If the p-value is significant (e.g., p < 0.05), we can confidently assert that the observed synchrony is greater than chance.

## 4. Peak Picking

While the heatmap is visually informative, we often want to extract the precise lags where coordination is strongest within each time window. The `pick_peaks()` function identifies local maximums within the WCC grid.

We specify the `L_size` argument to set the local search region.

```{r peak-picking}
# Extract peaks using a local search size of 5
wcc_peaks_df <- pick_peaks(wcc_results, L_size = 5)

print(wcc_peaks_df)
```

This returns a `wcc_peaks` data frame containing the elapsed time indices, the peak lags, and the corresponding correlation values.

## 5. Visualizing the Results

Finally, we can visualize the shifting synchronization landscape. The `plot_peaks_overlay()` function generates a heatmap of the correlations and plots the extracted peaks directly on top.

By passing the `time_step` argument, the axes are automatically converted from raw frame indices to seconds.

```{r visualization, fig.width=7, fig.height=5}
plot_peaks_overlay(
  wcc_obj = wcc_results,
  peaks_df = wcc_peaks_df,
  time_step = 1 / fs,
  show_zero_lag = TRUE
)
```

In the resulting plot, you should clearly see the structural shift we built into our data. For the first 30 seconds, the peak synchrony sits securely at a positive lag (indicating Person A leads). Right at the 30-second mark, the band of high correlation drops to a negative lag (indicating Person B is now leading).
