---
title: "WCC Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WCC Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Analyzing Interpersonal Synchrony with Windowed Cross-Correlation

This vignette walks through a complete Windowed Cross-Correlation (WCC) analysis using the `bsync` package. WCC is highly effective for quantifying interpersonal synchrony because it accommodates non-stationary relationships.  Unlike global cross-correlation, WCC captures how synchronization ebbs and flows over time.

We will cover data simulation, calculating WCC, surrogate testing for statistical significance, optima extraction, and quantifying leadership dynamics using the package's pipeline approach. For guidance on selecting the appropriate window sizes and lag parameters, please see the `suggest_wcc_params()` vignette.

## 1. Simulating and Preparing Realistic Data

To demonstrate the workflow, we will simulate a realistic interaction between two participants (Person A and Person B) captured at 30 Hz. We will generate smooth continuous data to simulate bodily motion and then add random measurement noise.

Unlike a perfectly stationary interaction, this scenario mimics natural turn-taking and includes four distinct behavioral phases over the course of 60 seconds:

1. **Phase 1 (0 to 20s):** Person A leads the interaction by 10 frames (0.33 seconds).
2. **Phase 2 (20 to 30s):** An uncorrelated lull where both participants move independently.
3. **Phase 3 (30 to 46s):** Person B takes over and leads by 12 frames (0.40 seconds).
4. **Phase 4 (46 to 60s):** Person A interjects and leads by 8 frames (0.26 seconds).

```{r setup-data, message=FALSE}
library(bsync)
library(dplyr)

set.seed(2026)

# Simulation parameters
fs <- 30
n_frames <- 1800 # 60 seconds of data

# 1. Generate underlying rhythms (a shared rhythm and two independent rhythms)
# We use a moving average to create smooth, human-like motion waves
time_seq <- seq(0, by = 1/fs, length.out = n_frames)
shared_base <- as.numeric(stats::filter(rnorm(n_frames + 300), rep(1/15, 15), circular = TRUE))
indep_base_A <- as.numeric(stats::filter(rnorm(n_frames + 300), rep(1/15, 15), circular = TRUE))
indep_base_B <- as.numeric(stats::filter(rnorm(n_frames + 300), rep(1/15, 15), circular = TRUE))

person_A_raw <- numeric(n_frames)
person_B_raw <- numeric(n_frames)

# 2. Piece together the dyadic interaction
for (i in 1:n_frames) {

  if (i <= 600) {
    # Phase 1 (0-20s): Person A leads by 10 frames (0.33 seconds)
    person_A_raw[i] <- shared_base[150 + i]
    person_B_raw[i] <- shared_base[150 + i - 10]

  } else if (i <= 900) {
    # Phase 2 (20-30s): Lull / No correlation
    person_A_raw[i] <- indep_base_A[150 + i]
    person_B_raw[i] <- indep_base_B[150 + i]

  } else if (i <= 1400) {
    # Phase 3 (30-46s): Person B leads by 12 frames (0.40 seconds)
    person_A_raw[i] <- shared_base[150 + i + 12]
    person_B_raw[i] <- shared_base[150 + i]

  } else {
    # Phase 4 (46-60s): Person A leads by 8 frames (0.26 seconds)
    person_A_raw[i] <- shared_base[150 + i]
    person_B_raw[i] <- shared_base[150 + i - 8]
  }
}

# 3. Create the raw data frame with added measurement noise
dyad_data_raw <- data.frame(
  time = time_seq,
  person_A_raw = person_A_raw + rnorm(n_frames, sd = 0.05),
  person_B_raw = person_B_raw + rnorm(n_frames, sd = 0.05)
)
```

### 1.1 Smoothing and Edge Trimming

Raw kinematic data almost always requires smoothing before analysis. Here, we apply a Savitzky-Golay filter to iron out the high-frequency measurement noise.

However, polynomial smoothing algorithms require surrounding data to calculate an accurate average. This causes mathematical distortion at the extreme beginning and end of the recording. We use `trim_edges()` to drop these edge artifacts to ensure our analysis is built strictly on stable data. A standard rule of thumb is to trim a length equal to your smoothing window.

```{r smooth-and-trim}
# 4. Smooth the raw data and trim the edges
dyad_data <-
  dyad_data_raw |>
  mutate(
    person_A = smooth_signal(person_A_raw, method = "sgolay", window = 15),
    person_B = smooth_signal(person_B_raw, method = "sgolay", window = 15)
  ) |>
  trim_edges(trim_length = 15)
```

## 2. Calculating Windowed Cross-Correlation

With our data ready, we can run the primary `wcc()` function.  This function slides a window across the time series and calculates the cross-correlation at various lags within each window.

We will use a window size of 90 frames (3 seconds) and a maximum lag of 45 frames (1.5 seconds). We will also increment the window by 30 frames (1 second) to allow for overlap and smooth transitions.

```{r calculate-wcc}
wcc_results <- wcc(
  x = dyad_data$person_A,
  y = dyad_data$person_B,
  time = dyad_data$time,
  window_size = 90,
  lag_max = 45,
  window_increment = 30,
  lag_increment = 1
)

# View a summary of the results
summary(wcc_results)
```

The `wcc()` function returns a list object of class `wcc_res` containing the results data frame, the overall Fisher's Z score, and the input settings.

We can easily visualize these initial results using the default `plot()` method. By passing the `time_step` argument, the axes are automatically converted from raw frame indices to seconds.

```{r plot-base-wcc, fig.width=7, fig.height=5}
plot(wcc_results, time_step = 1 / fs)
```

This generates a heatmap where deep blue indicates strong positive correlation (synchrony) and deep red indicates strong negative correlation. You can already see a clear track of high correlation shifting across the zero-lag line over time alongside a washed-out period of low correlation in the middle.

## 3. Surrogate Testing for Significance

Time series data are inherently autocorrelated. Because of this, high cross-correlation values can sometimes occur purely by chance. To test if the synchronization we observed is meaningful, we generate a null distribution using the circular shift method.

```{r surrogate-testing}
surrogate_results <- wcc_surrogate(
  x = dyad_data$person_A,
  y = dyad_data$person_B,
  time = dyad_data$time,
  window_size = 90,
  lag_max = 45,
  window_increment = 30,
  lag_increment = 1,
  n_surrogates = 1000
)

print(surrogate_results)
```

The output gives us an empirical p-value by calculating the proportion of surrogate Fisher's Z scores that meet or exceed our observed Fisher's Z. Because the observed value was higher than all 1000 permutations, the empirical p-value is reported as < .001.

## 4. Optima Extraction

While the heatmap generated above is visually informative, we often want to extract the precise lags where coordination is strongest within each time window. The `pick_optima()` function identifies local maximums within the WCC grid.

Because our simulation includes a programmed lull where the participants move independently, the algorithm will naturally find weak, meaningless "noise peaks" during that period. To prevent the tracking line from wildly jumping around during this lull, we can pass a `threshold` argument.

Setting `threshold = 0.25` ensures that any optimum with an absolute correlation weaker than r = 0.25 is overwritten with `NA`.

```{r optima-extraction}
# Extract optima using a local search size of 9 and a threshold of 0.25
wcc_optima_df <- pick_optima(wcc_results, L_size = 9, threshold = 0.25)

# View a statistical breakdown of the extracted optima
summary(wcc_optima_df)
```

### 4.1 Interpreting the Optima Summary

The `summary()` method provides a concise breakdown of the behavioral dynamics. We can map these results directly to the four phases we programmed into our simulation:

* **Completeness:** The summary will show that roughly 16% of the optima were dropped (set to `NA`). This perfectly aligns with Phase 2 of our simulation, which was a 10-second uncorrelated lull out of the 60-second total. The threshold successfully identified and removed this non-interactive period.
* **Lag Directionality (Leadership):** Because we supplied `person_A` as `x` and `person_B` as `y`, a positive lag means Person A is leading and a negative lag means Person B is leading. The summary will show positive lags for roughly 68% of the valid interaction (corresponding to the 34 total seconds Person A led in Phases 1 and 4). It will show negative lags for the remaining 32% (corresponding to the 16 seconds Person B led in Phase 3).
* **Optimum Value Distribution:** The quantiles provide a quick look at the strength of the coordination during the valid interactive phases, confirming that the retained peaks represent strong, structural synchrony.

### 4.2 Tuning the Local Search Window (`L_size`)

Choosing the right `L_size` is crucial for a successful local search. `L_size` must be an odd integer, and it defines the width of the neighborhood (in frames) that the algorithm evaluates to confirm a local maximum.

There is a delicate balance to strike:

* **If `L_size` is too small (e.g., 3):** The algorithm might get trapped by tiny noise ripples near lag zero and completely miss the true interactive response.
* **If `L_size` is too large (e.g., 45):** The algorithm might skip over the fast, immediate interactive response and lock onto a delayed, mathematically stronger, but theoretically irrelevant rhythm.

A good rule of thumb is to set `L_size` to capture roughly 0.1 to 0.5 seconds of data. At 30 Hz, an `L_size` between 5 and 15 frames is usually a great starting point. Because `pick_optima()` is computationally lightweight, we recommend experimenting with several values and visually comparing the plots.

## 5. Visualizing the Results

We can visually confirm our optima extraction by overlaying the tracking path directly onto the WCC heatmap.

```{r visualization, fig.width=7, fig.height=5}
plot_optima_overlay(
  surface_obj = wcc_results,
  optima_df = wcc_optima_df,
  time_step = 1 / fs,
  show_zero_lag = TRUE
)
```

In the resulting plot, the tracking line perfectly visualizes the distinct phases of our simulated interaction:

* **0 to 20 seconds:** A stable vertical line sits at a positive lag (~0.33s), correctly identifying Person A as the initial leader.
* **20 to 30 seconds:** There is a clean break in the tracking line. Because we applied a threshold, the algorithm successfully ignored the weak, random noise peaks during the uncorrelated lull, leaving this non-interactive period appropriately blank.
* **30 to 60 seconds:** The tracking line reappears and cleanly shifts to new lags, accurately capturing the transitions in leadership dynamics during the final two phases.

## 6. Quantifying Leadership Dynamics (The Pipeline Approach)

Visualizing the optima is helpful, but researchers ultimately need a continuous, quantifiable metric of who is driving the interaction.  The `leadership_asymmetry()` function converts the extracted optima into a bounded Leadership Asymmetry Index (LAI) ranging from -1 (y entirely leads) to 1 (x entirely leads).

Because `bsync` is designed around a consistent class structure, you can seamlessly chain the entire analytical process together using the native R pipe (`|>`):

```{r pipeline-lai, fig.width=7, fig.height=5}
# Run the complete pipeline from results to LAI visualization
wcc_results |>
  pick_optima() |>
  leadership_asymmetry(epoch_size = 10, min_valid = 3) |>
  plot(smooth = TRUE)
```

By grouping the windows into local epochs (in this case, groups of 10 windows), the LAI function smooths over momentary frame-by-frame jitter to reveal the broader structural periods of dominance. The resulting plot makes it immediately obvious when the conversation shifted from Person A driving the interaction to Person B taking over.
